Dashboard section

### 2022-07-26 Report/Draft Dashboard section.Rmd
Updating with info on report; updating some of the code as well.

>Ca. 6,5 sider.

@@

One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed. The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny.

@@

If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
```{r dashboard_KPI_calculation}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (revenue_private_contributions +
            revenue_membership_fees - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

@@

In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour of the box based on the calculated number. The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
```{r dashboard_valuebox, eval=FALSE}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

All the boxes also include a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1.
```{r dashboard_fontawesome, eval=FALSE}
# Using fontawesome for more up-to-date (and currently supported) fontawesome
# icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

@@

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots.

>>>>

>Ca. 8 sider.

@@

One of the outputs we have created is an example dashboard that calculates some KPIs and creates a few visualisations based on user-uploaded data from Proff.no and the Excel template we developed. The basic layout and structure of the dashboard were created using Flexdashboard, whilst most of the logic 'under the hood' was created using Shiny. Additionally, the dashboard provides the user with the option of downloading a pdf version of the analysis, using a parameterised R Markdown document.

@@

If there are any characters, the calculation outputs `NA`, otherwise it carries out the calculation. One example:
```{r dashboard_KPI_calculation, eval=FALSE}
# Creating Fundraising ROI value
KPI_fundraising_num <- reactive({
  # testing for data quality, returning NA if bad
  if (datatest() == TRUE) {
    NA
  } else {
    exceldata() |>
      # most recent year
      filter(year == max(year)) |>
      transmute(
        # Calculating the value as a percentage; removing all other
        KPI_fundraising_num =
          (sum(revenue_private_contributions,
            revenue_membership_fees,
            na.rm = TRUE
          ) - expenses_fundraising) /
            expenses_fundraising * 100
      ) |>
      # just want the single number
      pull() |>
      first()
  }
})
```

@@

In the case of some value boxes, such as the Fundraising ROI one, the function includes logic that changes the colour and icon of the box based on the calculated number. The levels are based on some cursory examination of expected levels of Fundraising ROI, but may of course be changed if the dashboard is developed further.
```{r dashboard_valuebox, eval=FALSE}
# Creating value box for fundraising ROI
output$KPI_box_fundraising <- renderValueBox({
  valueBox(
    value = KPI_to_string(KPI_fundraising_num()),
    caption = "Fundraising ROI",
    # Colour depends on value
    color = if (is.na(KPI_fundraising_num())) {
      "primary"
    } else if (KPI_fundraising_num() < 100) {
      "danger"
    } else if (KPI_fundraising_num() < 200) {
      "warning"
    } else {
      "success"
    },
    # icon depends on whether or not the value is NA
    icon = if (is.na(KPI_fundraising_num())) {
      "fa-solid fa-circle-xmark"
    } else {
      "fas fa-donate"
    }
  )
})
```

All the boxes also include, as mentioned, a Fontawesome icon, fetched using the fontawesome package, as Flexdashboard relies on the unsupported version 3 of Fontawesome, whilst fontawesome used the still-supported version 5 when coding began, and has just been updated to use version 6.1.1. All boxes use the theme colour 'primary' and the icon 'fa-solid fa-circle-xmark' if the value is NA.
```{r dashboard_fontawesome, eval=FALSE}
# Using fontawesome for more up-to-date (and currently supported) fontawesome
# icons
htmltools::tagList(fontawesome::fa_html_dependency())
```

@@

Finally, the dashboard contains a separate tab with some basic details about the KPIs and plots. At the top it includes a heading containing the name of the NGO as written in the Excel file:
```{r dashboard_title_creation, eval=FALSE}
# NGO title ==================================================================
NGO <- reactive({
  # validating the file again
  validate(need(
    tools::file_ext(input$excel$datapath) == "xlsx",
    "Please upload an Excel file."
  ))
  # fetching the name
  read_excel(input$excel$datapath) |>
    # Turning the needed column into a list
    pull(Firm) |>
    # discarding all NAs in case the name is somewhere
    purrr::discard(is.na) |>
    # choosing the first available name
    first()
})

title <- reactive({
  # testing if NGO() is NA
  if (is.na(NGO())) {
    # Printing warning if NA
    "Name of NGO missing"
    # Printing name of NGO
  } else {
    NGO()
  }
})
```

The title is output using `htmlOutput()` to use some basic html styling:
```{r dashboard_title, eval=FALSE}
# Creating NGO title using html formatting (colour based on MetBrewer
# "Hokusai2")
output$title <- renderText({
  paste('<b><font size="5em" color="#2f70a1">', title(), "</font></b>")
})
```

## The parameterised R Markdown report

The dashboard also allows the user to download a pdf version of the analysis through clicking on a button. A `downloadHandler()` function passes the necessary parameters to the R Markdown document and displays a temporary notification while the file is being created. The function includes temporary file creation arguments that should enable it to function on e.g. shinyapps.io and similar server contexts, but this has not been tested.
```{r}
## Download and parameters -----------------------------------------------------
output$report <- downloadHandler(

  # Based on code in Mastering Shiny
  filename = "report.pdf",
  content = function(file) {
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    # parameters
    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end
    )

    # notification
    id <- showNotification(
      "Generating report...",
      duration = NULL,
      closeButton = FALSE
    )
    on.exit(removeNotification(id), add = TRUE)

    # R Markdown render
    rmarkdown::render(tempReport,
      output_file = file,
      params = params,
      envir = new.env(parent = globalenv())
    )
  }
)
```

The code in the R Markdown document is very similar to that of the dashboard, but structured in a more conventional way. Data warnings are here text rather than blank plots, and the KPIs are integrated into the text instead of being presented in value boxes.
